---
title: "Intermediate-Elective"
author: "Madeline Hulle"
date: "2026-05-10"
format: pdf

---
# Set Up


## Packages 
```{r}
#| label: Packages
#| message: false

# general use and cleaning

library(tidyverse)
library(here)
library(janitor)
library(NatParksPalettes)


# reading in .xlsx files
library(readxl)

```

## Data

```{r}

birds <- read_csv(here("data", "birds.csv"))

aq_ins <- read_xlsx(
  here("data", "Aquatic Sampling Data-2026-03-10.xlsx"),
  sheet = "Aquatic Insects")

sites <- read_xlsx(
  here("data", "Aquatic Sampling Data-2026-03-10.xlsx"),
  sheet = "sites"
)
```
## Cleaning sites

```{r}
#| label: cleaning-sites

# creating new object from sites
ncos_sites <- sites |> 
  #filtering to only include NCOS quarterly sites
  filter(sector == "NCOS" & 
           sampling_frequency == "Quarterly")

```

## Cleaning aq_ins
```{r}
#| label: cleaning_aquatic_insects

# create a new object
aq_ins_clean <- aq_ins |> 
  # clean column names
  clean_names() |> 
  semi_join(ncos_sites, by = "site") |> 
  mutate(site_full = case_when(
    site == "NVBR" ~ "North of Venoco Bridge",
    site == "NEC" ~ "East Channel",
    site == "NMC" ~ "Main Channel",
    site == "NPB" ~ "South of Phelps Bridge",
    site == "NPB1" ~ "North of Phelps Bridge",
    site == "NPB2" ~ "Phelps Road",
    site == "NWP" ~  "West Pond",
    site == "NDC" ~ "Devereux Creek"))



```

# Problems 

## 1. Explain Your Inspiration

- I chose Cristina Cametti’s “IKEA most popular designers” visualization for my inspiration. 

- This made sense for the `birds` dataset because the ikea visualization was the comparison of two categorical variables, which `birds` has. The useage of designers in the ikea visualization inspired me to use informaion from a column which we havent used before - Data Enterer. 

- 










# TidyTuesday 03/11/2020 - IKEA Furniture

```{r}

# remove rows where either person or site is missing, group by site and person
dt <- aq_ins_clean |> 
  filter(!is.na(person_entering_data), !is.na(site_full)) |> 
  group_by(site_full, person_entering_data) |> 
# count entries per person/site combination
  tally() |> 
  ungroup() |> 
# remove low-count entries to reduce crowding
  filter(n >= 2) |>
  arrange(desc(n)) |> 
# convert to factors
  mutate(site_full = as.factor(site_full),
         person_entering_data = as.factor(person_entering_data))

# number of blank bars to add as spacing
empty_bar <- 2

# create empty rows for spacing
to_add <- matrix(NA, empty_bar, ncol(dt))
# match column names to dt
colnames(to_add) <- colnames(dt)
# append empty rows to dt
dt <- rbind(dt, to_add)

# give each bar a numeric position on the x-axis
dt$id <- seq(1, nrow(dt))

# copy dt for label calculations
label_data <- dt
# count total number of bars
number_of_bar <- nrow(label_data)

# calculate rotation angle for each label based on its position in the circle
angle <- 90 - 360 * (label_data$id - 0.5) / number_of_bar

# align labels on the left side of the circle
label_data$hjust <- ifelse(angle < -90, 1, 0)
# flip labels on the left side so they stay readable
label_data$angle <- ifelse(angle < -90, angle + 180, angle)

# initialize plot with bars colored by site
p <- ggplot(dt) +
  geom_bar(aes(x = as.factor(id), y = n, fill = site_full), stat = "identity") +
# negative lower bound creates hollow center, upper bound gives room above tallest bar
  ylim(-10, max(dt$n, na.rm = TRUE) + 10) +
  theme_minimal() +
# apply custom color palette
  scale_fill_manual(values = natparks.pals("DeathValley")) +
  theme(
# remove axis labels, titles, and grid lines
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1, 4), "cm")
  ) +
# add title in the center of the wheel
  annotate(geom = "text",
           x = 0, y = -10,
           hjust = .5, vjust = 1,
           label = "NCOS Aquatic Invertebrate\nSamples by Data Enterer",
           size = 2.5, lineheight = .8,
           color = "black") +
# wrap bar chart into a circle
  coord_polar(start = 0) +
# add person name labels rotated to follow the circle
  geom_text(data = label_data,
            aes(x = id, y = n + 1,
                label = person_entering_data,
                hjust = hjust),
            color = "black", fontface = "bold", alpha = 0.6, size = 3,
            angle = label_data$angle,
            inherit.aes = FALSE)
p

```




