---
title: "Intermediate-Elective"
author: "Madeline Hulle"
date: "2026-05-10"
format: pdf

---
# Set Up


## Packages 
```{r}
#| label: Packages
#| message: false

# general use and cleaning

library(tidyverse)
library(here)
library(janitor)
library(NatParksPalettes)


# reading in .xlsx files
library(readxl)

```

## Data

```{r}


aq_ins <- read_xlsx(
  here("data", "Aquatic Sampling Data-2026-03-10.xlsx"),
  sheet = "Aquatic Insects")

```


## Cleaning aq_ins
```{r}
#| label: cleaning_aquatic_insects

# create a new object
aq_ins_clean <- aq_ins |>
  # only including columns of interest
  select(`Person Entering Data`, Site) |> 
  # clean column names
  clean_names() |> 
  semi_join(ncos_sites, by = "site") |> 
  mutate(site_full = case_when(
    site == "NVBR" ~ "North of Venoco Bridge",
    site == "NEC" ~ "East Channel",
    site == "NMC" ~ "Main Channel",
    site == "NPB" ~ "South of Phelps Bridge",
    site == "NPB1" ~ "North of Phelps Bridge",
    site == "NPB2" ~ "Phelps Road",
    site == "NWP" ~  "West Pond",
    site == "NDC" ~ "Devereux Creek"))



```

# Problems 

## 1. Explain Your Inspiration

- I chose Cristina Cametti’s “IKEA most popular designers” visualization for my inspiration. 

- This made sense for the `aq_ins_clean` dataset because the ikea visualization was the comparison of two categorical variables, which `aq_ins_clean` has. The usage of designers in the ikea visualization inspired me to use information from a column which we haven used before - Data Enterer. 

- The variables being shown from the `aq_ins_clean` dataset are `person_entering_data` and `site_full`. The person entering data will be represented by one bar and the colors making up the bar (and the relative sizes of these colors) will represent the site that observation came from and the number of samples entered. 

## 2. 








# TidyTuesday 03/11/2020 - IKEA Furniture

```{r}

# wrangle: one row per person/site combo, with person totals for sorting
dt <- aq_ins_clean |> 
  filter(!is.na(person_entering_data), !is.na(site_full)) |> 
  group_by(person_entering_data, site_full) |> 
  tally() |> 
  ungroup() |> 
  group_by(person_entering_data) |> 
  mutate(total = sum(n)) |> 
  ungroup() |> 
  filter(total >= 2) |> 
  arrange(desc(total)) |> 
  mutate(person_entering_data = fct_reorder(person_entering_data, total, .desc = TRUE),
         site_full = as.factor(site_full))

# one row per person for angle/label calculations
label_data <- dt |> 
  group_by(person_entering_data) |> 
  summarize(total = first(total)) |> 
  arrange(desc(total)) |> 
  mutate(id = row_number())

number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id - 0.5) / number_of_bar
label_data$hjust <- ifelse(angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle + 180, angle)

# plot
p <- ggplot(dt) +
  geom_bar(aes(x = person_entering_data, y = n, fill = site_full),
           stat = "identity", position = "stack") +
  ylim(-10, max(label_data$total, na.rm = TRUE) + 5) +
  theme_minimal() +
  scale_fill_manual(values = natparks.pals("DeathValley"), name = "Site") +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1, 4), "cm")
  ) +
  annotate(geom = "text",
           x = 0, y = -10,
           hjust = .5, vjust = 1,
           label = "NCOS Aquatic Invertebrate\nSamples by Data Enterer",
           size = 2.4, lineheight = .8,
           color = "black") +
  coord_polar(start = 0) +
  geom_text(data = label_data,
            aes(x = person_entering_data, y = total + 1,
                label = person_entering_data,
                hjust = hjust),
            color = "black", fontface = "bold", alpha = 0.6, size = 3,
            angle = label_data$angle,
            inherit.aes = FALSE)
p

```




